Developing Decarbonisation Pathways for Irish homes in changing TIMES

### 2021-08-23 manuscript.tex
23 Aug

%PICKUPHERE
\section{Results}
\label{Results}

The results from the newly developed TIM provide some valuable insights on each sector and the sectoral interconnections. The results outlined in this paper are based upon six scenarios, which are focused around the 2030 target. While TIM can only analysis the energy system, the other large source of GHG emissions is from agriculture. Therefore, some scenarios are labelled based on agriculture 2030 GHG reduction percentage and energy system 2030 GHG reduction percentage, below is a brief description of the six scenarios:
\begin{enumerate}[(a)]
    \item \textbf{No Mitigation} Least cost solution, where no emissions constraints are imposed
    \item \textbf{With Additional Measures (WAM)} includes the implementation of Ireland’s CAP 2019 towards 2030 and net-zero in 2050.
    \item \textbf{A25E65} Agriculture GHG emissions reduce by 25\%, energy system GHG emissions reduces by 65\% by 2030 compared to 2018 \& net-zero in 2050 
    \item \textbf{A33E61} Agriculture GHGs reduces by 33\% and Energy system GHGs by 61\% \& net-zero in 2050 
    \item \textbf{A40E57} Agriculture GHGs reduces by 40\% and energy system GHGs by 57\% \& net-zero in 2050 
    \item \textbf{A51E51} Equal GHG reduction share in of 51\% in Agriculture and energy Systems \& net-zero in 2050 
\end{enumerate}

While the first two scenarios do not comply with the Climate Action and Low Carbon Development (Amendment) Act 2021 target of an overall 51\% GHG reduction in 2030 compared to 2018, they do provide insights on the incremental steps required to achieve these targets. The last four scenarios comply with 2030 and 2050 targets and offer insights about different trajectories to achieve those targets.

\subsection{Residential Results}
\label{Results:Residential }
The residential sector reduces $CO_2$ emissions by between 63-71\% in 2030 for the four scenarios which comply with the Climate Bill. Compare this to the ``No Mitigation" scenario, which shows an increase of 4\% $CO_2$ in 2030 and the  ``WAM" scenario which shows a 8\% reduction in $CO_2$ by 2030. These results assist in representing the scale of decarbonisation required in the residential sector.

Focusing on the residential fuel-mix gives further insights on the optimal pathways to achieve 2030 targets. Figure \ref{fig:ResFuel} shows this, with the base year being represented on the top row, the four Climate Bill compliant scenarios on rows 2-5, and the last two rows are not compliant with the Climate Bill.

\begin{figure}[!htbp]
 \centering
 \includegraphics[width=\textwidth]{Figures/RSD_Fuel2.png}
 \caption{Residential Fuel per scenario}
 \label{fig:ResFuel}
\end{figure}

In 2018, the residential sector had 124 PJ of primary energy, which is needed to decrease by 8-13\% in 2030, to comply with the Climate Bill. Results show a large uptake of electricity in the residential sector within Climate Bill compliant scenarios. The large increase in ambient energy indicates the type of technology deployed in the residential sector, 
the overall share of ambient heat increases from 1.4\% in 2018 to between 13-26\% in 2030, more on this in Figure \ref{fig:HeatPump}. In Climate Bill compliant scenarios, peat and coal are completely phased out, while oil use decreases 90-93\% by 2030. Gas energy increase by 34\% in the A51E51 scenario, but shows little change in other Climate Bill compliant scenarios in 2030.

Figure \ref{fig:Retrofits} shows both the number of retrofits per type and the energy savings up to 2030, for each of the four Climate Bill compliant scenarios.

\begin{figure}[!htbp]
 \centering
 \includegraphics[width=\textwidth]{Figures/Retrofits.png}
 \caption{Retrofits by type per scenario}
 \label{fig:Retrofits}
\end{figure}

The number of retrofits varies from 393,000 in A51E51 to 615,000 in A25E65, meaning between 23-36\% of existing dwellings will require some form of retrofit.  
As mentioned in Table \ref{tab:HeatPumpConstraint}, the rollout of retrofits constraints the amount of heat pumps which can be installed, however, a certain share of existing dwellings and all new dwellings are heat pump ready, i.e B2 rating or above.

@@

@@

\section{Discussion}
\label{Discussion}

The methodology applied to the residential sector in TIM is transparent, up-to-date, and robust. The previous version of TIM, namely, the Ireland TIMES model, which was used for almost 10 years, and was continuously tweaked to react to the ever changing climate policy, climate targets, and technological advancements. 
\par TIM is developed to provide insights on sectoral carbon budgets and 2030 and 2050 climate targets. As the first iteration of TIM, the core model is expected to be improved, while on-going research will continue to  analyse certain sectors in greater detail. 
The large increase in electricity ?demand? may require some soft-linking with PLEXOS to investigate power generation operating constraints and reliability of supply. While LEAP is another powerful tool within the Energy Policy and Modelling Group of MaREI, which can be used to analyse individual climate policies.
\par The model will require some future developments in the residential sector, DH, and renewable gas. The large growth of data centres expected in Ireland can provide waste heat to DH networks. While the capacity of biomethane in Ireland can effect the potential decarbonisation of natural gas, adding biomethane to the natural gas grid has not been accounted for in the model.
\par The dissagreation of archetypes by BER ratings will add to the model complexity, but help provide insights into optimal BER ratings to target in retrofitting and heat pump installation. While the dissageration of the model into 26 regions will provide spatial insights. The data in TIM is setup to allow for this, but due to the complexity the model was simplified for the initial release.

>>>>

\section{Results}
\label{Results}

The results from TIM can provide some valuable insights on the residential sector and the sectoral interconnections when targeting carbon budgets, and long-term climate targets. While the purpose of this paper was to describe the residential methodology, the results used here only scrape the surface in terms of exploring residential decarbonisation. However, the results provide some initial insights into residential decarbonisation and are discussed in section \ref{Discussion}. Three different scenarios are used to explore the results, those scenarios and descriptions are:

\begin{enumerate}[(a)]
    \item \textbf{No Mitigation} Least cost solution, where no emissions constraints are imposed
    \item \textbf{Previous Effort } complies with Climate Action Plan 2019, reduces Non-ETS GHG by 30\% in 2030 compared to 2005, and net-zero in 2050.
    \item \textbf{Current Effort} complies with Climate Action Bill 2021, energy system GHG emissions reduces by 65\% by 2030 compared to 2018 \& net-zero in 2050 ( complies if agriculture GHG emissions reduces by 25\% in 2030 ).
\end{enumerate}

While the first two scenarios do not comply with the Climate Action and Low Carbon Development (Amendment) Act 2021 target of an overall 51\% GHG reduction in 2030 compared to 2018, they do provide insights on the incremental steps required to achieve these targets.

\subsection{Residential Results}
\label{Results:Residential }

The quantity and mix of fuel going into the residential for each scenario and year is  shown in Figure \ref{fig:ResFuel}. The base year is shown on the left, while annual snapshots of the fuel-mix for the other scenarios are shown for 2030, 2040, and 2050.

\begin{figure}[!htbp]
 \centering
 \includegraphics[width=\textwidth]{Figures/residential_fuelmix.png}
 \caption{Residential fuel-mix per scenario per year}
 \label{fig:ResFuel}
\end{figure}

In 2018, the residential sector required 124 PJ of primary energy, the increase in households, and energy efficiency improvements resulted in a 9\% primary energy increase in 2030, under the Previous Effort scenario. Whereas under the Current Effort scenario, the primary energy is required to reduce by 13\% in 2030, this highlights the large incremental step in ambition required when moving from Climate Action Plan 2019 to the Climate Action Bill 2021. In 2050, under the Previous Effort scenario, the total primary energy shows a continued growth 17\% higher than 2018, whereas the Current Effort scenario shows an overall reduction of 11\% which indicates a post-2030 growth. 
The fluctuation in the fuel-mix required is significant, especially to comply with 2030 targets, for example, oil reduces 60\% in the Previous Effort and 90\% in the Current Effort scenario in 2030. Furthermore, peat and coal are phased out in the residential sector in both scenarios. The lesser carbon intensive natural gas increases 211\% in Previous Effort without the upper constraint of the gas infrastructure and decreases 4\% in Current Effort scenario. The electricity continues to decarbonise rapidly in all scenarios, resulting in an increase in electricity in the residential sector. In both scenarios, the carbon intensity of electricity ($gCO_2/kWh$) falls by two-thirds, with the Current Effort scenario only marginally lower than the Previous Effort scenario. New emerging fuels which are prevalent in the Current Effort include space ambient heat and water ambient heat due to the increase of heat pumps in the residential sector, more on this in Figure \ref{fig:HeatPump}.

The optimal number of retrofits by type and quantity up to 2030, 2040 and 2050 are summarised in Figure \ref{fig:Retrofits}.

\begin{figure}[!htbp]
 \centering
 \includegraphics[width=\textwidth]{Figures/NumberofRetrofits.png}
 \caption{Retrofits by type per scenario}
 \label{fig:Retrofits}
\end{figure}

While the Climate Action Plan 2019 targeted 500,000 retrofits, the model only retrofitted 90,000 up to 2030, however, all these retrofits are deep and on larger detached dwellings. The number of retrofits jumps up to 615,000 in the Current Effort scenario, there is a mix of types of depth of retrofits with different dwelling types, but no apartments are retrofitted. The cumulative energy savings are displayed on the secondary axis.

As mentioned in Table \ref{tab:HeatPumpConstraint}, the rollout of retrofits constraints the amount of heat pumps which can be installed, however, a certain share of existing dwellings and all new dwellings are heat pump ready, i.e B2 rating or above.

@@

The residential sector reduces $CO_2$ emissions by between 63-71\% in 2030 for the four scenarios which comply with the Climate Bill. Compare this to the ``No Mitigation" scenario, which shows an increase of 4\% $CO_2$ in 2030 and the  ``WAM" scenario which shows a 8\% reduction in $CO_2$ by 2030. These results assist in representing the scale of decarbonisation required in the residential sector.

@@

\section{Discussion}
\label{Discussion}
SOLAR??
The methodology applied to the residential sector in TIM is transparent, up-to-date, and robust. The previous version of TIM, namely, the Ireland TIMES model, which was used for almost 10 years, and was continuously tweaked to react to the ever changing climate policy, climate targets, and technological advancements. 
\par TIM is developed to provide insights on sectoral carbon budgets and 2030 and 2050 climate targets. As the first iteration of TIM, the core model is expected to be improved, while on-going research will continue to  analyse certain sectors in greater detail. 
The large increase in electricity ?demand? may require some soft-linking with PLEXOS to investigate power generation operating constraints and reliability of supply. While LEAP is another powerful tool within the Energy Policy and Modelling Group of MaREI, which can be used to analyse individual climate policies.
\par The model will require some future developments in the residential sector, DH, and renewable gas. The large growth of data centres expected in Ireland can provide waste heat to DH networks. While the capacity of biomethane in Ireland can effect the potential decarbonisation of natural gas, adding biomethane to the natural gas grid has not been accounted for in the model.
\par The dissagreation of archetypes by BER ratings will add to the model complexity, but help provide insights into optimal BER ratings to target in retrofitting and heat pump installation. While the dissageration of the model into 26 regions will provide spatial insights. The data in TIM is setup to allow for this, but due to the complexity the model was simplified for the initial release.
